Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species

### 2019-12-25 Initial_Analysis.Rmd
Included RFMO analysis and missing values for results

# Set Jurisdiction to EEZ or RFMO
# Jurisdiction <- "EEZ"
Jurisdiction <- "RFMO"

@@

Clean_sau_eez_fes <- sau_eez_fes %>% 
  gather("Type","fishing_region",2,4) %>% 
  left_join(sau_eez_fes,
            by = ("fishing_entity_id")
  ) %>% 
  group_by(eez_name,fishing_entity) %>% 
  summarise(n()) %>% 
  ungroup() %>% 
  ### Manually Fix names for estimating catch
  mutate(eez_name = ifelse(eez_name =="Greece (without Crete)","Greece",
                           ifelse(eez_name =="Greenland (Denmark)","Greenland",
                                  ifelse(eez_name =="United Kingdom (UK)","United Kingdom",
                                         ifelse(fishing_entity =="Italy","Italy",
                                                ifelse(eez_name =="Chile (mainland)","Chile",
                                                       ifelse(fishing_entity =="Brazil","Brazil",
                                                              ifelse(eez_name == "Ecuador (mainland)", "Ecuador",
                                                                     ifelse(eez_name =="Portugal (mainland)","Portugal",
                                                                            ifelse(fishing_entity == "Korea (North)","Korea (North)",
                                                                                   paste(eez_name))
                                                                            )
                                                                     )
                                                              )))))),
         fishing_entity = ifelse(fishing_entity == "Viet Nam","Vietnam",fishing_entity
         )
  )

@@

# Read Sau Value Data (Thank you Tim!)
SAU_Revenue_Data <- fread(paste(Data_Path,"Distribution_Data/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>% 
  clean_names() %>% 
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise(mean_value = mean(value,na.rm=T))

# Read Species Status Data (Sub-routine Species Status)
Status_Data <- fread(paste(Results_Path,"Species_Status.csv", sep ="")) %>% 
  clean_names() %>% 
  rename(eez_name=territory)
# Status_Data <- Species_Status

@@

# Load and clean transboundary results
# Results_Trans <- GetTransResults(Spp)

# Clean results susing function CleanResultsfun
Clean_Results_Trans <- CleanResults(Results_Trans) %>% 
  clean_names() %>% 
  rename(eez_name = territory,
         fishing_entity = ct_fishing_entity,
         eez_neighbour = neighbour) %>% 
  mutate(fishing_entity = ifelse(eez_name == "Spain (Mediterranean and Gulf of Cadiz)","Spain",paste(fishing_entity)))

# Load and clean Discrete results
# Non_Transboundary_Data <- fread(paste(Results_Path,"Non_Transboundary_Data.csv",sep="")) %>% 
#   filter(!is.na(TaxonKey)) %>% 
#   select(2:3)

# Vulnerabillity Data from William
cc_vul <- fread(paste(Data_Path,"Species_Data/vulspp310817.csv", sep ="")) %>% 
  rename(taxon_key = V1,
         ccvul = V2)

fish_vul <- fread(paste(Data_Path,"Species_Data/FishSenIndex.csv", sep ="")) %>% 
  clean_names()

### RFMO species list

RFMO_Species <- read.csv(paste(Data_Path,"Spatial_Data/RFMO/clean_rfmos.csv", sep ="")) %>% 
  clean_names() %>% 
  left_join(Exploited_Species,
            by = c("taxon_name"))

# The path
path_world <- paste(Data_Path,"/Spatial_Data/","RFMO","/SAU_RFMO",sep="")
# #The File
fnam_world <- "SAU_RFMO_Sept_2015.shx"
# #Load it!
# 
RFMOs_sf <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world)) %>% 
  dplyr::rename(Name = RFMO_nm_s) %>% 
  st_set_crs(3832) %>% # Making sure is same proyection
  st_transform(3832) # to match shapefiles

Clean_Results_Trans <- Clean_Results_Trans %>% 
  filter(taxon_key %in% RFMO_Species$taxon_key)

@@

#### Determine transboundary and non-transboundary species and nations

@@

# Get Catch data for mapping (Supplemental materials)

Catch_Country <- SAU_Catch_Data %>%
  group_by(eez_name) %>%
  summarise(eez_catch = round(sum(total_catch,na.rm=T),2)) %>%
  left_join(Clean_sau_eez_fes,
            by = "eez_name") %>%
  group_by(fishing_entity) %>%
  summarise(fishing_entity_catch = sum(eez_catch,na.rm=T)/1000000) %>% # Million tones
  mutate(fishing_entity = ifelse(fishing_entity == "USA","United States of America",
                                 ifelse(fishing_entity == "Russian Federation","Russia",fishing_entity)),
         Bins = cut(fishing_entity_catch,breaks = 5),
         Bins_Plot = cut(fishing_entity_catch,breaks = 5,
                         labels = c('super low', 'low','median','high','super high'))
  ) %>% 
  filter(fishing_entity_catch > 0)

@@

```{r Territory_Map_Round, eval =F, echo =F}

@@

@@

```{r Territory_Map, eval =F, echo =F}

lenght(unique(Clean_Results_Transboundary$taxon_key))

Transboundary_Spp %>% 
  ggplot() +
  geom_bar(
    aes(
      x = reorder(eez_name,-n_trans_spp),
      y = n_trans_spp
    ),
    stat = "identity"
  ) +
  coord_flip()

@@

# Transboundary Species shapefile
World_sf_si <- RFMOs_sf %>%
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  # st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%
  left_join(Transboundary_Spp, by ="eez_name")

# ICCAT <-World_sf_si %>% 
#   filter(eez_name == "ICCAT")
#

@@

# Make map
Plot <- ggplot() +
  geom_sf(data = World_sf_si_land) +
  geom_sf(data = World_sf_si,
          aes(
            fill =  n_trans_spp#,
            # colour = reorder(Bins_Plot,n_trans_spp)
          ), 
          size = 0.1
          ) + # n_trans_spp or Bins
  # geom_sf_label(data = World_sf_si,
  #               aes(label = eez_name)
  #               ) +
  scale_fill_gradientn("Number of Shared Species",
    colours = wes_palette("Zissou1", 100, type = "continuous")) +
  # scale_fill_manual("",values = c("#5BBCD6",#"#5BBCD6", # blue
  #                                 "#00A08A",#"#00A08A", #green
  #                                 "#F2AD00",#"#F2AD00", # Yellow
  #                                 "#F98400",#"#F98400", #Orange
  #                                 "#FF0000"#,"#FF0000" #Red
  # ),
  # na.value = "grey"
  # ) +
  # scale_colour_manual("",values = c("#5BBCD6",#"#5BBCD6", # blue
  #                                   "#00A08A",#"#00A08A", #green
  #                                   "#F2AD00",#"#F2AD00", # Yellow
  #                                   "#F98400",#"#F98400", #Orange
  #                                   "#FF0000"#,"#FF0000" #Red
  # ),
  # na.value = "grey"
  # ) +
  ggtheme_map()

Plot_Name <- paste("RFMO_100_25.png", sep = "")

ggsave(Plot_Name,
       plot = Plot,
       width = 16,
       height = 8,
       units = "in"#,
       # path = Figures_Path
       )

@@

### Mackrell war

@@

Conflict_nations <- c()

Macrell <- Tresholded_T_Data %>% 
  filter(taxon_key == 600118,
         eez_name %in% Conflict_nations
  )

@@

Species_Data <- Tresholded_T_Data %>% 
  group_by(taxon_key) %>% 
  summarise(n_trans_spp = length(unique(fishing_entity))) %>% 
  left_join(Exploited_Species) %>%
  rename(scientific_name = taxon_name) %>% 
  left_join(Species_environ_list) %>%
  mutate(DemersPelag = ifelse(is.na(DemersPelag),"Other",
                              ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bento-pelagic",paste(DemersPelag)
                                     )
                              ),
         n_trans_spp = ifelse(n_trans_spp > 100, 100, n_trans_spp),
         DemersPelag = str_to_title(DemersPelag)
         ) %>% 
  select(individual=scientific_name,
         value = n_trans_spp,
         group = DemersPelag) %>% 
  arrange(group,value) %>% 
  filter(value > 20)  # remove all species under 20 (way too many to plot!)

#### Exosystem based plots

@@

@@

    # plot.margin = unit(c(-3.5, -3.5, -4, -4), "cm") #top, right, bottom, left
    plot.margin = unit(rep(-3.5,4),"cm")

@@

>>>>

# Set Jurisdiction to EEZ or RFMO
Jurisdiction <- "EEZ"
# Jurisdiction <- "RFMO"

@@

Clean_sau_eez_fes <- sau_eez_fes %>% 
  gather("Type","fishing_region",2,4) %>% 
  left_join(sau_eez_fes,
            by = ("fishing_entity_id")
  ) %>% 
  group_by(eez_name,fishing_entity,fishing_entity_id) %>% 
  summarise(n()) %>% 
  ungroup() %>% 
  ### Manually Fix names for estimating catch
  mutate(eez_name = ifelse(eez_name =="Greece (without Crete)","Greece",
                           ifelse(eez_name =="Greenland (Denmark)","Greenland",
                                  ifelse(eez_name =="United Kingdom (UK)","United Kingdom",
                                         ifelse(fishing_entity =="Italy","Italy",
                                                ifelse(eez_name =="Chile (mainland)","Chile",
                                                       ifelse(fishing_entity =="Brazil","Brazil",
                                                              ifelse(eez_name == "Ecuador (mainland)", "Ecuador",
                                                                     ifelse(eez_name =="Portugal (mainland)","Portugal",
                                                                            ifelse(fishing_entity == "Korea (North)","Korea (North)",
                                                                                   paste(eez_name))
                                                                     )
                                                              )
                                                       )))))),
         fishing_entity = ifelse(fishing_entity == "Viet Nam","Vietnam",fishing_entity
         )
  )

@@

# Read Sau Value Data (Thank you Tim!)
SAU_Revenue_Data <- fread(paste(Data_Path,"Distribution_Data/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>% 
  clean_names() %>% 
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise(mean_value = mean(value,na.rm=T)) %>% 
  left_join(Clean_sau_eez_fes)

# Read Species Status Data (Sub-routine Species Status)
Status_Data <- fread(paste(Results_Path,"Species_Status.csv", sep ="")) %>% 
  clean_names() %>% 
  rename(eez_name=territory)

@@

# Read Results
Clean_Results_Trans <- GetTransResults(Spp) %>% 
  CleanResults(.) %>% 
  clean_names() %>% 
  rename(eez_name = territory,
         fishing_entity = ct_fishing_entity,
         eez_neighbour = neighbour) %>% 
  mutate(fishing_entity = ifelse(eez_name == "Spain (Mediterranean and Gulf of Cadiz)","Spain",paste(fishing_entity)))

# Vulnerabillity Data from William
# cc_vul <- fread(paste(Data_Path,"Species_Data/vulspp310817.csv", sep ="")) %>% 
#   rename(taxon_key = V1,
#          ccvul = V2)

# fish_vul <- fread(paste(Data_Path,"Species_Data/FishSenIndex.csv", sep ="")) %>% 
  # clean_names()

# ---------------------------- #
# RFMOs Results ####
# ---------------------------- #

# List of species
Spp_rfmo <- list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Results/RFMO/Trans_Results/")

# List of RMOs
rfmo_list <- read.csv("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Spatial_Data/RFMO/clean_rfmos.csv") %>% 
  clean_names() %>% 
  left_join(Exploited_Species) %>% 
  filter(!rfmo_name %in% c("IPHC","PSC","NASCO")) # remove these

# Load and clean transboundary results
Results_rfmo <- GetTransResults(Spp_rfmo) %>% 
  clean_names() %>% 
  rename(rfmo_name = territory) %>% 
  # left_join(rfmo_list,
  #           by =c("rfmo_name","taxon_key"))
  filter(
    taxon_key %in% rfmo_list$taxon_key, # just the taxa managed by the rfmos
    !rfmo_name %in% c("IPHC","PSC","NASCO"), # remove these
    !neighbour %in% c("IPHC","PSC","NASCO") # remove these
  )

# The path
path_world <- paste(Data_Path,"/Spatial_Data/","RFMO","/SAU_RFMO",sep="")
# #The File
fnam_world <- "SAU_RFMO_Sept_2015.shx"
# #Load it!
# 
RFMOs_sf <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world)) %>% 
  dplyr::rename(Name = RFMO_nm_s)

@@

#### Determine transboundary and species and nations

@@

@@

```{r Territory_Map_Round_Revenue, eval =F, echo =F}

@@

```{r Territory_Map_Round_Catch, eval =F, echo =F}

# Get Catch data for mapping (Supplemental materials)
Catch_Breaks <- c(-1,10,100,1000,5000,10000)

Catch_Country <- SAU_Catch_Data %>%
  group_by(eez_name) %>%
  summarise(eez_catch = round(sum(total_catch,na.rm=T),2)) %>%
  left_join(Clean_sau_eez_fes,
            by = "eez_name") %>%
  group_by(fishing_entity) %>%
  summarise(fishing_entity_catch = sum(eez_catch,na.rm=T)/1000) %>% # thousand tones
  mutate(
    # fishing_entity = ifelse(fishing_entity == "USA","United States of America",
  #                                ifelse(fishing_entity == "Russian Federation","Russia",fishing_entity)),
         Bins = cut(fishing_entity_catch,breaks = Catch_Breaks),
         Bins_Plot = cut(fishing_entity_catch,breaks = Catch_Breaks,
                         labels = c('super low', 'low','median','high','super high'))
  ) %>% 
  filter(fishing_entity_catch > 0)

Bins_Catch <- Catch_Country %>% 
  arrange(Bins) %>% 
  pull(Bins) %>% 
  unique()

Labels_Trans <- c("(1-29]",paste(Bins_Trans[2:5]))
Labels_Catch <- c("(1-10]",paste(Bins_Catch[2:5]))

# Transboundary Species shapefile
World_sf_si <- World_sf %>%
  st_transform(crs = 4326) %>%
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%
  left_join(Transboundary_Spp, by ="eez_name")

# Total catch/Revenue shapefile
World_sf_si_land <- World_land_sf %>%
  st_transform(crs = 4326) %>%
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
  left_join(Catch_Country, by ="fishing_entity")

# Get the NA's to be grey
Grey_land <- World_sf_si_land %>% 
  filter(is.na(Bins_Plot))

# Make map
Map <- ggplot() +
  geom_sf(data = World_sf_si_land,aes(fill = reorder(Bins_Plot,fishing_entity_catch)), colour = "grey80",size = 0.1) +
  geom_sf(data = World_sf_si,aes(fill =  reorder(Bins_Plot,n_trans_spp)),colour = "grey20", size = 0.1) + # n_trans_spp or Bins
  geom_sf(data = Grey_land,aes(fill = Bins),fill = "grey80", colour = "grey50",size = 0.1) + # n_trans_spp or Bins
  scale_fill_manual("",values = c("#5BBCD6",#"#5BBCD6", # blue
                                  "#00A08A",#"#00A08A", #green
                                  "#F2AD00",#"#F2AD00", # Yellow
                                  "#F98400",#"#F98400", #Orange
                                  "#FF0000"#,"#FF0000" #Red
  ),
  na.translate = F
  ) +
  ggtheme_map() +
  # Transboundary labels
  annotate("text",
           x =(c(-85.5,-30,25,81,136)+5),
           y = -93.5,
           label = Labels_Trans,
           color="black",
           size=3,
           angle=0,
           # fontface="bold",
           hjust=0) +
  # Catch labels
  annotate("text",
           x =(c(-85.5,-31,20,75,131)+5),
           y = -110,
           label = Labels_Catch,
           color="black",
           size=3,
           angle=0,
           # fontface="bold",
           hjust=0) +
  # # Legend Titles
  annotate("text",
           x =-180,
           y = -93.5,
           label = "Transboundary Spp. (n)",
           color="black",
           size=3,
           angle=0,
           fontface="bold",
           hjust=0) +
  annotate("text",
           x =-180,
           y = -110,
           label = "Catch (Thousand tons)",
           color="black",
           size=3,
           angle=0,
           fontface="bold",
           hjust=0) +
  theme(
    legend.box.margin = margin(t = -54, l = 130),
    legend.text = element_blank(),
  ) +
  guides(fill = guide_legend(nrow = 1))

ggsave(
    plot = last_plot(),
    width = 8,
    height = 5,
    units = "in",
    filename = "S1_Catch_EEZ.png"#,
    # path = Figures_Path
  )

@@

```{r RFMO_Map, eval =F, echo =F}

Tresholded_rfmo_Data <- Results_rfmo %>% 
  filter(
    area_index >= Min_AI,
    area_index <= Max_AI,
    model_index >= MI
  ) %>% 
  group_by(rfmo_name) %>% 
  summarise(
    n_trans_spp = length(unique(taxon_key)),
    n_neighbours = length(unique(neighbour)),
    trans_Rrate = n_trans_spp/n_neighbours
  )

# Number of RFMOs
rfmo_list %>% 
  pull(rfmo_name) %>% 
  unique() %>% 
  length() # 15

# Number of total taxa
rfmo_list %>% 
  pull(taxon_key) %>% 
  unique() %>% 
  length() # 99

# Number of taxa per RFMO
rfmo_list %>% 
  group_by(rfmo_name) %>% 
  summarise(in_trans_spp = length(unique(taxon_key)))

# Number of shared taxa
Results_rfmo %>% 
  pull(taxon_key) %>% 
  unique() %>% 
  length() # 81

### Bar plot

Tresholded_rfmo_Data %>% 
  ggplot() +
  geom_bar(
    aes(
      x = reorder(rfmo_name,-n_trans_spp),
      y = n_trans_spp
    ),
    stat = "identity"
  ) +
  coord_flip()

@@

# Transboundary Species shapefile
World_sf_si <- RFMOs_sf %>%
  filter(!Name %in% c("IPHC","PSC","NASCO")) %>% 
  rename(rfmo_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  # st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%
  left_join(Tresholded_rfmo_Data, by ="rfmo_name")

@@

# Make map
Plot <- ggplot() +
  geom_sf(data = World_sf_si,
          aes(
            fill =  n_trans_spp,
            colour = n_trans_spp
          ), 
          size = 0.1
          ) + # n_trans_spp or Bins
  geom_sf(data = World_sf_si_land) +
  # geom_sf_label(data = World_sf_si,
  #               aes(label = eez_name)
  #               ) +
  scale_fill_gradientn("Number of Shared Species",
    colours = wes_palette("Zissou1", 100, type = "continuous"),
    limits = c(0,45),
    breaks = seq(0,45,5)
    ) +
  scale_colour_gradientn("Number of Shared Species",
    colours = wes_palette("Zissou1", 100, type = "continuous"),
    limits = c(0,45),
    breaks = seq(0,45,5)
    ) +
  ggtheme_map()

Plot_Name <- paste("S2_RFMO.png", sep = "")

ggsave(Plot_Name,
       plot = Plot,
       width = 16,
       height = 8,
       units = "in",
       path = Figures_Path
       )
```

@@

### Mackerel (Scomber scombrus) war

@@

Exploited_Species %>% 
  filter(taxon_name == "Scomber scombrus") # 600118

NW_Europe <- UN_Regions %>% 
  filter(sub_region  %in% c("Western Europe", "Northern Europe")) %>% 
  pull(fishing_entity) %>% 
  unique()

Conflict_nations <- c(NW_Europe,"Portugal","Spain")

Mackerel_Rev <- SAU_Revenue_Data %>% 
  filter(taxon_key == 600118,
         fishing_entity %in% Conflict_nations
         ) %>% 
  group_by() %>% 
  summarise(sum(mean_value))/1000000

Mackerel_Catch <- SAU_Catch_Data %>% 
  left_join(Clean_sau_eez_fes) %>% 
  filter(taxon_key == 600118,
         fishing_entity %in% Conflict_nations
         ) %>% 
  group_by() %>% 
  summarise(sum(total_catch))/1000

@@

Species_Data <- Tresholded_T_Data %>% 
  group_by(taxon_key) %>% 
  summarise(n_trans_spp = length(unique(fishing_entity))) %>% 
  left_join(Exploited_Species) %>%
  rename(scientific_name = taxon_name) %>% 
  left_join(Species_environ_list) %>%
  mutate(DemersPelag = ifelse(is.na(DemersPelag),"Other",
                              ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bento-pelagic",paste(DemersPelag)
                                     )
                              ),
         n_trans_spp = ifelse(n_trans_spp > 100, 100, n_trans_spp),
         DemersPelag = str_to_title(DemersPelag)
         ) %>% 
  select(individual=scientific_name,
         value = n_trans_spp,
         group = DemersPelag) %>% 
  arrange(group,value)  %>% 
  mutate(
    individual = ifelse(individual == "Chloroscombrus chrysurus", "C. chrysurus",
                        ifelse(individual == "Scomberomorus commerson", "S. commerson",
                               individual)
    )
  )

#### Ecosystem based plots

@@

Species_Data <- Species_Data %>%
  filter(value > 20)

@@

    plot.margin = unit(rep(-3.5,4.5),"cm")

@@

## RFMO specifics

```{r RFMOs, eval = T, echo = F}

### Results exploration

ICCAT_List <- rfmo_list %>% 
  filter(rfmo_name == "ICCAT") %>% 
  pull(taxon_key) %>% 
  unique()

Tresholded_T_Data %>% 
  filter(taxon_key %in% ICCAT_List) %>% 
  group_by(taxon_key) %>% 
  summarise(
    n_countries = length(unique(fishing_entity))
  )

ICCAT_Trans <- Species_Data %>% 
  rename(taxon_name = individual) %>% 
  left_join(Exploited_Species) %>% 
  filter(taxon_key %in% ICCAT_List)

### 2019-12-25 Manuscript_Draft.Rmd
Included RFMO analysis and missing values for results

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These 5 nations are responsable for 30% of the global revenue from transboundary species.

Catches, however, are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@SOFIA2018], getting around 1/3 of its capture from transboundary species. For the US, 16 species make up 90% of the revenue from transboudary species, with American scallop (*Placopecten magellanicus*) shared with Canada in Georges Bank [@https://elischolar.library.yale.edu/cgi/viewcontent.cgi?article=1008&context=fes-pubs] contributing 41.8% or the revenue. Russia shares the largest fished species in the world, Alaskan polllock (*Theragra chalcogramma*) [@SOFIA2018], with both Japan and the US. This fishery is responsible for 50% of the Russian revenue from transboundary species, with other eight species comprising 90% of the revenue. In the case of Peru, Peruvian anchoveta (*Engraulis ringens*) and jumbo squid (*Dosidicus gigas*), both shared with Chile, represent almost 90% of the revenue from transboundary species, with anchoveta accounting for 79%.

@@

While these catch trends do not directly reflect stock status [@Branch:2011br] - e.g. other factors like markets, fishing policies or extreem events also influence catch volume [@PaulyNature2013] - they could ressemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Empirical analysis suggest that management of transboundary species will have better outcomes in terms of fish yield and sustainabillity whith cooperation between nations [@REF]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw].

Collaboration can also reduce conflict between nations over shared ressources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Miller2003]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for mackerel [@Spijkers:2017ij], a species that is transboundary and contributes **X** to landings and **X** in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky].

@@

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@NelsonFishesOfTheWorld]. Not surprisingly, highly migratory pelagic-oceanic species are the only group that significantly differs in terms of sharing countries (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 nations sharing these resources (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are catalogued as stradling species by FAO definitions and globally managed by Regional Fisheries Management Organizations (RFMOs) [@FAO]. Sixteen major RFMOs in the world oversee the management of 101 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 56.4% of these species (*n* = 57) are actually shared by neighbouring RFMOs with the North Atlantic Salmon Conservation Organization (NASCO), the North East Atlantic Fisheries Commission (NASCO) and the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 30 species with neighbouring RFMOs (Suplemental figure RFMO_MAP).

The catch from these species comes mainly from five South Pacific Islands nations, from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (PNA) (Fig. 3B). The PNA manages 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year [@https://www.ffa.int/nauru_agreement]. However, the effectiveness of RFMOs responsible for managing high migratory stocks, has been previously questioned [@CullisSuzuki:2010fi].

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these might be cosmopolitan but not necessarily highly migratory (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va].

Moreover, many coastal communities, specially in developing regions, and highly dependent on fishing local fish (e.g. reef fish) for food security and improved livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf; @https://www.sciencedirect.com/science/article/pii/S0308597X17308357]. For example, from the top five countries fishing transboundary reef - associated species, Indonesia, Malaysia and the Philippines are among the higest fish consumption per capita in the world with 20, 60 and 40 kg per year, respectiveley [@https://www.sciencedirect.com/science/article/pii/S0308597X12001315]. In the Phillipines...

@@

<!-- # Recicled text (Not to be included) -->

@@

>>>>

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These 5 nations are responsable for 30% of the global revenue from transboundary species. Peru and Russia lead the production of transboundary species harvesting over 5,000 thousand tones (**Suplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the (currently) largest fishery in the world, Alaskan polllock (*Theragra chalcogramma*) with both Japan and the US [@SOFIA2018]. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@SOFIA2018] getting around 1/3 of its capture from transboundary species, followde by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

@@

While these catch trends do not directly reflect stock status [@Branch:2011br] - e.g. other factors like markets, fishing policies or extreem events also influence catch volume [@PaulyNature2013] - they could ressemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Empirical analysis suggest that management of transboundary species will have better outcomes in terms of fish yield and sustainabillity whith cooperation between nations [@REF]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Collaboration can also reduce conflict between nations over shared ressources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Miller2003]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky].

@@

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@NelsonFishesOfTheWorld]. Not surprisingly, highly migratory pelagic-oceanic species are the only group that significantly differs in terms of sharing countries (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 nations shared per species (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are catalogued as stradling species by FAO definitions and globally managed by Regional Fisheries Management Organizations (RFMOs) [@FAO]. Fifteen major RFMOs in the world oversee the management of 99 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species (*n* = 81) are actually shared within neighbouring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighbouring RFMOs (**Suplemental figure RFMO_MAP**). The catch from pelagic-oceanic species comes mainly from five South Pacific Islands nations, from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (PNA) (Fig. 3B). The PNA manages 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year [@https://www.ffa.int/nauru_agreement].

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these might be cosmopolitan but not necessarily highly migratory (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Moreover, many coastal communities, specially in developing regions, and highly dependent on fishing local fish (e.g. reef fish) for food security and improved livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf; @https://www.sciencedirect.com/science/article/pii/S0308597X17308357]. For example, from the top five countries fishing transboundary reef - associated species, Indonesia, Malaysia and the Philippines are among the higest fish consumption per capita in the world with 20, 60 and 40 kg per year, respectiveley [@https://www.sciencedirect.com/science/article/pii/S0308597X12001315].

@@

```{r Call_Sup_1, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species per EEZ and they contribution to countries' EEZ catch. The number of species is represented in the EEZ while the catch (thousand tonnes) is represented in the land polygon."}

img <- readPNG(here("./Figures/S1_Catch_EEZ.png"))
grid.raster(img)
```

![**Number of transboundary species per EEZ and they contribution to countries' EEZ catch. The number of species is represented in the EEZ while the catch (thousand tonnes) is represented in the land polygon. **](./Figures/S1_Catch_EEZ.png)

```{r Call_Sup_2, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of shared species by RFMOs. Analysis included with 85 species managed by Fourteen RFMOs"}

img <- readPNG(here("./Figures/S2_RFMO.png"))
grid.raster(img)
```

![**Number of shared species by RFMOs. Analysis included with 99 species managed by Fourteen RFMOs **](./Figures/S2_RFMO.png)

**Suplemental figure Trans_Catch_MAP**

# Recicled text (Not to be included)

@@

<!-- However, the effectiveness of RFMOs responsible for managing high migratory stocks, has been previously questioned [@CullisSuzuki:2010fi]. -->
